### 2019-04-15 paper/draft_esa.tex
Fix ESA draft Latex compilation

\documentclass[a4paper,UKenglish,cleveref, autoref]{lipics-v2019}

\bibliographystyle{plainurl}% the mandatory bibstyle

\input{my_preamble.tex}

\title{Metric Spaces with Expensive Distances}
\titlerunning{Metric Spaces with Expensive Distances}
% uncomment for LIPICS 
\author{Anonymous Authors}{Anon}{}{}{}
%\author{Michael Kerber}{TU Graz}{kerber@tugraz.at}{https://orcid.org/0000-0002-8030-9299}{Supported by Austrian Science Fund (FWF) grant number P 29984-N35}
%\author{Arnur Nigmetov}{TU Graz}{nigmetov@tugraz.at}{}{}
%\authorrunning{M. Kerber and A. Nigmetov}
\authorrunning{Anonymous Authors}

\keywords{metric spaces, doubling dimension, spanners, approximate nearest neighbor}
\ccsdesc[500]{Theory of computation~Random search heuristics}
\ccsdesc[500]{Theory of computation~Random search heuristics}
%% \category{dummy}

% \category{dummy}

%\Copyright{Michael Kerber and Arnur Nigmetov}
\Copyright{Anonymous Authors}
% END uncomment for LIPICS

\def \expDistFigPath {pics/}
\def \expDistDataPath {./}

\begin{document}
\maketitle

\begin{abstract}
In algorithms for finite metric spaces, it is common
to assume that the distance between two points
can be computed in constant time, and complexity
bounds are expressed only in terms of the number
of points of the metric space.
We introduce a different model where we assume
that the computation of a single distance is
an expensive operation and consequently, the goal is
to minimize the number of such distance queries.
This model is motivated by metric spaces
that appear in the context of topological data analysis.

We consider two standard operations on metric spaces,
namely the construction of a $1+\eps$-spanner
and the computation of an approximate nearest
neighbor for a given query point.
In both cases, we partially explore the metric space
through distance queries and infer lower and upper bounds
for yet unexplored distances through triangle inequality.
For spanners, we evaluate several exploration strategies
through extensive experimental evaluation.
For approximate nearest neighbors, we prove that our
strategy returns an approximate nearest neighbor
after a logarithmic number of distance queries.
\end{abstract}

\input{expensive-distances-main.tex}

\bibliography{bib}

\end{document}